---
title: "data_collection"
author: "Sujan Bhattarai"
date: "2024-02-02"
output: html_document
---

```{r}
# load the packages
library(dplyr)
library(RSQLite)
library(ggplot2)
```

```{r}
# create a connection to the database
soccer_dataset <- dbConnect(RSQLite::SQLite(), "data/soccer_data/database.sqlite")

# list all tables
dbListTables(soccer_dataset)
```

```{r}
#create dataframe for each of this different tables from this database

Country <- dbGetQuery(soccer_dataset, "SELECT * FROM Country")
League  <- dbGetQuery(soccer_dataset, "SELECT * FROM League")
Match   <- dbGetQuery(soccer_dataset, "SELECT * FROM Match") 
Player  <- dbGetQuery(soccer_dataset, "SELECT * FROM Player")
Player_Attributes <- dbGetQuery(soccer_dataset, "SELECT * FROM Player_Attributes")
Team <- dbGetQuery(soccer_dataset, "SELECT * FROM Team")
Team_Attributes <- dbGetQuery(soccer_dataset, "SELECT * FROM Team_Attributes")
```

```{r}
# join country and league dataframes based on primary key country_id
country_league <- inner_join(Country, League, by = c("id" = "country_id"), suffix = c("_country", "_league"))

# join the country_league dataframe with the match dataframe based on primary key league_id and country_id
country_league_match <- inner_join(country_league, Match, 
                                   by = c("id_league" = "league_id", "id" = "country_id"), 
                                   suffix = c("_country_league", "_match"))

```
There is no primary key associated with country_league_match dataframe that I can use to join with player attributes. So, I will be considering these two dataframes separetely, and make visualization and analysis based on these two dataframes.

```{r}
# combine player and player_attributes dataframes based on primary key player_api_id
player_player_attributes <- inner_join(Player, Player_Attributes, 
                                       by = c("player_api_id" = "player_api_id"), 
                                       suffix = c("_player", "_player_attributes"))
```

Data Visualization steps: 
 - Distribution of shot power among players

```{r}
# plot the overall distribution of shot power among players
ggplot(player_player_attributes, aes(x = shot_power)) +
  geom_histogram(binwidth = 5, fill = "#E84D8A") +
  theme_minimal() +
  xlab("Shot power") +
  ylab("Frequency") +
  theme(axis.text = element_text(size = 10))

```

- Top 20 players with excellent shot power

```{r}
# player who has excellent shooting skills, i.e. shot_power > 90. The number 90 is entirely personal preference
excellent_shot_power <- player_player_attributes %>%
  filter(shot_power > 90) %>%
  select(player_name, shot_power, overall_rating) %>% 
  distinct(player_name, .keep_all = TRUE) %>% 
  arrange(desc(shot_power)) %>% 
  #filter top 20 players with excellent shot power
  head(20)

```

-Visualization of the top 20 players with their names

```{r}
#reorder the player_name in descending order of shot_power
excellent_shot_power$player_name <- factor(excellent_shot_power$player_name, levels = excellent_shot_power$player_name)

#plot the players with excellent shot power
ggplot(excellent_shot_power, aes(y = player_name, x = shot_power)) +
  # plot with lollipop
  geom_segment(aes(x = 0, xend = shot_power, y = player_name, yend = player_name), color = "black") +
  # add points
  geom_point(shape = 20, fill = "black", color = "#E84D8A", size = 3) +
  # use minimal theme that has white background
  theme_minimal() +
  # rotate the y-axis text, set hjust to 1 to align 
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 1)) +
  # remove the axis title for yaxis
  ylab("") +
  xlab("Shot power") +
  # remove the grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  # set the cartesian between 85 and 95 to highlight this area
  coord_cartesian(xlim = c(85, 95)) +
  # add the text label for shot power and put it to the left of the point
  geom_text(aes(label = shot_power), hjust = -0.5, size = 2) +
  #zoom into the area of interest
  scale_x_continuous(breaks = seq(85, 95, 5), labels = function(x) ifelse(x == 85, 0, x)) +
  #increase font size
  theme(axis.text = element_text(size = 10))

```

## does these top 20 players with excellent shot power has good, bad or the ugly overall ratings ?
```{r, warning=FALSE}
# read the player_player_attributes dataset, craete scatterplot and plot size based on overall rating
ggplot(player_player_attributes, aes(x = shot_power, y = overall_rating)) +
  geom_point(color = "#E84D8A", alpha = 0.01) +
  theme_minimal() +
  xlab("Shot power") +
  ylab("Overall rating") +
  theme(axis.text = element_text(size = 10)) +
  scale_size_continuous(range = c(1, 10)) + 
  #fit a linear model to the data
  geom_smooth(method = "lm", se = FALSE, color = "#7f58AF")

```

```{r}

```



